---
title: "Regression"
author: "Claire Herdeman"
date: "11/18/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Binary Variables and Regression

To illustrate the binary variable case, generate fake data the represents sex and height

```{r df}
# Made dataframe
df <- data_frame(female=c(0,0,0,1,1,1), height=c(152,140,168,120,136,125))
df

# Find average height by group
df %>% group_by(female) %>% summarise_all(mean)
```


Now run a regression of our binary variable on height: 

```{r reg}
# Run a linear regression
r <- lm(height ~ female, df)
rs <- summary(r)
# Show summary of regression output
rs
```

## Plot it

```{r plot}
df %>% ggplot(aes(female, height)) + 
  geom_point() + 
  geom_abline(intercept = rs$coefficients['(Intercept)','Estimate'], 
              slope = rs$coefficients['female','Estimate'])
```

## Testing
H_0: The means of the two groups are equal
H_A: They are not equal

In the binary case, that means that we need to check whether the difference between the two group means is 0, i.e. whether $\beta_1$ = 0.

The t-test is reported in the output table, how is it calculated?
```{r tests}
beta1_hat <- rs$coefficients['female','Estimate']
se_beta1_hat <- rs$coefficients['female','Std. Error']
ttest <- beta1_hat/se_beta1_hat
ttest
```

Note that in this case the F-test is $t^2$ (test it and compare with the summary regression output).