---
title: "Roy Model"
output: html_document
---

### Monday, October 8 5:00 - 6:20 PM

## Setup

The section loads libraries that we will need to use to run the code below.

```{r setup}
# For this session we will load the tidyverse, a commonly used set of R libraries
# Find more information here: https://www.tidyverse.org/packages/
library('tidyverse')
library('MASS')
```

## Roy Model

In the context of Prof. Black's lecture, the Roy Model was used to motivate the need for economic and statistical models to underly our analysis of observed phenomena in public policy. To (hopefully) illustrate the model as well as some of the power of computational tools, here is a brief simulation of the model. The expectation is NOT that you can put this together yourself right now.

### Generate Random Data

The first step is to generate random data from two correlated distributions with the characteristics described in class, i.e. Economist earnings are normally distributed with mean = 60k and stdev = 10k, and Accountant earnings are normally distributed with mean = 65k and st dev = 5k. This generation will produce data that pairs draws from both distribution, i.e. we are observing for every theoretical individual in our sample both their potential earnings as an accountant and their potential earnings as an economist. 

```{r data}
set.seed(10042018)
samples <- 30000 # Sample size or size of population
cor <- 0.86 # correlation coefficient

# Generate correlated std normal random sample using the mvnorm function from MASS package
data <- mvrnorm(n=samples, mu=c(0, 0), Sigma=matrix(c(1, cor, cor, 1), nrow=2), empirical=TRUE)

# Convert to dataframe
df1 <- dplyr::as_data_frame(data)

# Make the dataframe easier to use 
df1 <- df1 %>% dplyr::rename(accnt = V1, econ = V2) # rename columns

# Change distributions from std normal to those specified in lecture:
# Accounting ~ N(65000, 5000)
# Economics ~ N(60000, 10000)
mu_econ <- 60000
sigma_econ <- 10000
mu_accnt <- 65000
sigma_accnt <- 5000

df1 <- df1 %>% mutate(
        accnt = accnt*sigma_accnt + mu_accnt, # update accounting variable
        econ = econ*sigma_econ + mu_econ # update econ variable
)
```

### Sanity Checks

Let's run a few checks to make sure data looks reasonable. First we'll use the `head` and `tail` commands to get a look at the data. Then, we will make sure that the correlation between the accoutant and economist distributions is 0.86, the value we set when we generated the data, using the `cor` command. Finally, we will run the `summarise` command to see the mean, min, max, and quartiles of each of the distributions.

```{r sanity}
# Look at the first and last 6 rows of the dataframe
head(df1)
tail(df1)

# Compare the correlation we set to the correlation we calculate
check <- round(cor(df1$accnt, df1$econ)) == round(cor)
print(ifelse(check, "The correlations are the same!", "Oops, the correlations are not the same."))

# Generate summary statistics 
print("Economists Summary Stats")
summary(df1$econ)
print("Accountant Summary Stats")
summary(df1$accnt)
```

All of these results are in line with what we expect, so let's move on to visualizing the distributions using ggplot2.

```{r plot}
# Visualize the distributions
df1 %>% 
    gather(key = id, value = c('econ', 'accnt')) %>% # reshapes the data to a 'long' format
    dplyr::rename(value = 'c("econ", "accnt")') %>% # rename the column for ease of use
    ggplot(aes(x=value, color=id)) + # call the ggplot function
        geom_density(fill="white", alpha=0.5, position="identity") # make a density plot
```
```{r plot2}
# Contour plot overlayed on heat map image of results
df1 %>% 
  ggplot(aes(x = accnt, y = econ)) +
  geom_density2d() +      
  #scale_alpha_continuous(limits=c(0,1000),breaks=seq(0,1000,by=1000)) +
  geom_point(colour="red",alpha=0.01)

bivn.kde <- kde2d(df1$accnt/1000, df1$econ/1000, n = 50)   # from MASS package
image(bivn.kde)
contour(bivn.kde, add = TRUE)
```
