---
title: "Hypothesis Testing"
author: "Claire Herdeman"
date: "11/12/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem Statement
We have a "gambling machine" that we believe gives positive results 25% of the time, i.e. p = 0.25. Using the binomial distribution, we can calculate the probability that each number of successes, 0 through 32, occurs. We will also calculate for p = 0.5.

```{r data}
n = 32
p1 = 0.25
p2 = 0.5

gambling <- data_frame(num = 0:32,
           p25 = dbinom(num, n, p1),
           p50 = dbinom(num, n, p2),
           c25 = pbinom(num, n, p1),
           c50 = pbinom(num, n, p2))

head(gambling)
tail(gambling)
```


## Plot the distribution
Let see what the pmfs look like for the different parameters p.

```{r pmf}
gpmf <- gambling %>% 
  ggplot() +
  geom_point(aes(x = num, y = p25), color='blue') +
  geom_point(aes(x = num, y = p50), color='red') 

gpmf
```

```{r cdf}
gcdf <- gambling %>% 
  ggplot() +
  geom_point(aes(x = num, y = c25), color='blue') +
  geom_point(aes(x = num, y = c50), color='red') 

gcdf
```

## Define significance level
Significance levels are typically defined at 5%. We are asking: what is the probability that we would see the observed data under the null hpyothesis? We think that if that probability is less than the significance level (i.e. 5%), 
```{r Ha_less}
sig <- 0.05

# Test Ha: p < 0.25
gcdf + geom_hline(yintercept = sig)
```

```{r Ha_greater}
# Test Ha: p > 0.25
gcdf + geom_hline(yintercept = 1 - sig)
```

```{r Ha_neq}
# Test Ha: p != 0.25
gcdf + geom_hline(yintercept = sig/2) + geom_hline(yintercept = 1 - sig/2)
```