---
title: "Discrete Random Variables"
author: "Claire Herdeman"
date: "10/6/2018"
output: html_document
---
# TA Session 1
### Monday, October 8 5:00 - 6:20 PM

## Setup

The section loads libraries that we will need to use to run the code below.

```{r setup}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Probability Mass Function

First, generate some random data from a normal distribution (mean = 0, sd = 1). Round so that our variable is discrete rather than continuous.

```{r gen}
set.seed(10062018)
samples = 1000
dist <- as_data_frame(x = round(rnorm(samples, 0, 1)))
head(dist)
```

Now we count the freqency with which each value occurs, then calculate the probability that draw each value.

Does this satisfy the properties of the PMF?
What is f(-3)? f(0)?

```{r count}
prob <- dist %>% 
          count(value) %>% # count how many times each value in the x column occurs
          mutate(pr = n/samples) # divide the count by sample size to get probability

prob
```

We can also show the PMF graphically. 

```{r graph}
dist %>% 
  ggplot(aes(x = value)) +
  geom_histogram(binwidth = 1, color = 'black', fill = 'white', alpha = 0.5)  
```

## Cumulative Distribution Function

We can start from the prob dataframe that we defined above. We will define a new cum column (for cumulative) that is the cumulative sum of all preceding probabilities.

Does this satisfy the properties of the CDF?
What is F(-3)? F(0)?

```{r cdf}
prob <- prob %>%
          mutate(cum = cumsum(pr))

prob
```

Let's take a graphical look.

```{r cdf_graph}
dist %>%
  ggplot(aes(x = value)) +
  stat_ecdf()
```