---
title: "Roy Model"
output: html_document
---

### Monday, October 8 5:00 - 6:20 PM

## Setup

The section loads libraries that we will need to use to run the code below.

```{r setup}
# For this session we will load the tidyverse, a commonly used set of R libraries
# Find more information here: https://www.tidyverse.org/packages/
library('tidyverse')
library('MASS')
```

## Roy Model

In the context of Prof. Black's lecture, the Roy Model was used to motivate the need for economic and statistical models to underly our analysis of observed phenomena in public policy. To (hopefully) illustrate the model as well as some of the power of computational tools, here is a brief simulation of the model. The expectation is NOT that you can put this together yourself right now.

### Generate Random Data

The first step is to generate random data from two correlated distributions with the characteristics described in class, i.e. Economist earnings are normally distributed with mean = 60k and stdev = 10k, and Accountant earnings are normally distributed with mean = 65k and st dev = 5k. This generation will produce data that pairs draws from both distribution, i.e. we are observing for every theoretical individual in our sample both their potential earnings as an accountant and their potential earnings as an economist. 

```{r data}
set.seed(10042018)
samples <- 100000 # Sample size or size of population
cor <- 0.86 # correlation coefficient

# Generate correlated std normal random sample using the mvnorm function from MASS package
data <- mvrnorm(n=samples, mu=c(0, 0), Sigma=matrix(c(1, cor, cor, 1), nrow=2), empirical=TRUE)

# Convert to dataframe
df1 <- as_data_frame(data)

# Make the dataframe easier to use 
df1 <- df1 %>% dplyr::rename(accnt = V1, econ = V2) # rename columns

# Change distributions from std normal to those specified in lecture:
# Accounting ~ N(65000, 5000)
# Economics ~ N(60000, 10000)
mu_econ <- 60000
sigma_econ <- 10000
mu_accnt <- 65000
sigma_accnt <- 5000

df1 <- df1 %>% mutate(
        accnt = accnt*sigma_accnt + mu_accnt, # update accounting variable
        econ = econ*sigma_econ + mu_econ # update econ variable
)
```

### Sanity Checks

Let's run a few checks to make sure data looks reasonable. First we'll use the `head` and `tail` commands to get a look at the data. Then, we will make sure that the correlation between the accoutant and economist distributions is 0.86, the value we set when we generated the data, using the `cor` command. Finally, we will run the `summarise` command to see the mean, min, max, and quartiles of each of the distributions.

```{r sanity}
# Look at the first and last 6 rows of the dataframe
head(df1)
tail(df1)

# Compare the correlation we set to the correlation we calculate
check <- round(cor(df1$accnt, df1$econ)) == round(cor)
print(ifelse(check, "The correlations are the same!", "Oops, the correlations are not the same."))

# Generate summary statistics 
print("Economists Summary Stats")
summary(df1$econ)
print("Accountant Summary Stats")
summary(df1$accnt)
```

All of these results are in line with what we expect, so let's move on to visualizing the distributions using ggplot2.

```{r plot}
# Visualize the distributions
df1 %>% 
    gather(key = id, value = c('econ', 'accnt')) %>% # reshapes the data to a 'long' format
    dplyr::rename(value = 'c("econ", "accnt")') %>% # rename the column for ease of use
    ggplot(aes(x=value, color=id)) + # call the ggplot function
        geom_density(fill="white", alpha=0.5, position="identity") # make a density plot
```

The density graph above clearly shows what Prof. Black higlighted in class - that though the MEAN of the accountant's salary is higher than that of the economists, the economists distribution is broader, i.e. has fatter tails, meaning that if you are on the high end of the distribution, you are more likely to earn a higher salary as an economist. We know that these distributions are highly correlated (think: the skillset to be good at economics is similar, though not exactly the same as the skillset to be good at accounting), meaning that if you are at the high end of one distribution, you are likely to be at the high end of the other. 

### Picking a Career

Our assumption is that every person will choose to be an accountant or an economist based on what will maximize their salary. We will assign job labels strictly based on where an individual will earn more, i.e. Y<sub>i</sub> = max(Y<sub>e,i</sub>, Y<sub>a,i</sub>)

```{r jobs}
# Assign job labels
df1 <- df1 %>% mutate(job = ifelse(econ > accnt, "econ", "accnt"))

# Take a look at the change to the dataframe using head
head(df1)
```

### Results

```{r results}
# Make results dataframe
results <- df1 %>% 
            group_by(job) %>% # Group all of the rows with the same "job" together
            summarise('Economist Earnings' = mean(econ), 'Accountant Earnings' = mean(accnt),
                      n=n()) %>% # Calculate means and counts for economicts and accountants
            mutate(job = c("Accountant", "Economist")) %>% # add labels
            t() # transpose

colnames(results) <- c('Accountant', 'Economist')

results <- as.data.frame(results) %>%
            slice(2:4) %>%
            mutate(
              x= c('Economist Earnings', 'Accountant Earnings', 'n')
            ) %>%
            dplyr::select(x, Accountant, Economist)

results
```

The columns are the job, and the rows are potential earnings. So in this case the (Economicst, Economist Earnings) cell is the average salary of economists that we observe in the real world. The (Economist, Accounting Earnings) is a COUNTERFACTUAL that is unobserved in the world, i.e. the average salary of economist had they become accountants instead. As discussed in lecture, a naive observer would simply would look at the observed salaries in (i.e. (Economicst, Economist Earnings) and (Accountant, Accountant Earnings)) and think that everyone should be an economist. However, the Roy model helps us understand that individuals are making rational choices about their employment based on characteristics that aren't easily observable. This is how economic and statistical models can help us make sense of what we see in the world.

For a final check, calculate the weighted average of our observed and counterfactual outcomes (i.e., economists who became economists and accountants if they had become economists) and see if they come close to the distribution means we set initially. That is, close to 65k for accountants and 60k for economists. Spoiler: it works!
