---
title: "Bootstrap"
author: "Claire Herdeman"
date: "12/3/2018"
output: pdf_document
---

```{r setup, warning=FALSE}
library(tidyverse)
```

```{r load}
df <- read.csv("~/Downloads/ChicagoMSAWomen.csv")
levels(df$uhrswork)[83] <- 99
levels(df$uhrswork)[84] <- 0
df$uhrswork <- as.numeric(as.character(df$uhrswork))
set.seed(12032018)
```

## Estimate $\hat{F}$

With bootstrap simulations, we are estimating the distribution of a variable in the population using out sample. Using the Chicago MSA data, we will estimate the distribution of the hours worked (uhrswork) variable. 
```{r cdf}
df <- df %>% arrange(uhrswork)
N <- length(df$uhrswork)
dist <- df %>% 
  count(uhrswork) %>% 
  mutate(prop = n/N,
         cdf = cumsum(prop))

head(dist)

dist %>% ggplot() +
  geom_col(aes(x = uhrswork, y = prop), fill = "blue") +
  geom_step(aes(x = uhrswork, y = cdf)) +
  xlim(0, 100) +
  scale_x_continuous(breaks=c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100))

```

## Sampling from the data
To generate bootstrap estimates, we sample from the $\hat{F}$ distribution, which can be done in R using the sample command.
```{r cars}
# Set the sample size
n <- 10
# Note that bootstrap estimates require resampling, so must have replace = T 
sample(df$uhrswork, size = n, replace = TRUE)
```

## Bootstrap Draws
To generate a boostrap distribution, pick B, the number of bootstrap resamples and sample from the data.
```{r pressure, results="hide"}
B = 10000 # number of bootstrap replications

# initialize empty df
df10 <- data_frame(obs=1:n)

for (i in 1:B) {
  print(i)
  cola <- paste('col', i, sep= '_')
  df10[[cola]] <- sample(df$uhrswork, size = n, replace = TRUE)
}
```

Summarise each for the parameter of interest (in this case the mean)
```{r summarise10}
# Summarise each sample
means10 <- df10 %>% 
  select(contains("_")) %>% 
  summarise_all(mean) %>%
  t %>%
  as.data.frame

head(means10,10)

# Plot distribution
means10 %>% ggplot(aes(x=V1)) +
  geom_density(color='black') +
  xlim(20, 70)

# Bootstrap estimate of the mean:
bootstrap10 <- means10 %>% summarise_all(funs(mean))
```

```{r n_1000, results="hide"}
n <- 1000
# initialize empty df
df1000 <- data_frame(obs=1:n)

for (i in 1:B) {
  print(i)
  cola <- paste('col', i, sep= '_')
  df1000[[cola]] <- sample(df$uhrswork, size = n, replace = TRUE)
}
```

```{r summarise1000}
# Summarise each sample
means1000 <- df1000 %>% 
  select(contains("_")) %>% 
  summarise_all(mean) %>%
  t %>%
  as.data.frame

head(means1000,10)

# Plot distribution
means1000 %>% ggplot(aes(x=V1)) +
  geom_density(color='black') +
  xlim(20, 70)

bootstrap1000 <- means1000 %>% summarise_all(funs(mean))
```


```{r compare}
mean(df$uhrswork)
bootstrap10
bootstrap1000
```

```{r n_all, results="hide"}
n <- length(df$uhrswork)
# initialize empty df
dflength <- data_frame(obs=1:n)

for (i in 1:B) {
  print(i)
  cola <- paste('col', i, sep= '_')
  dflength[[cola]] <- sample(df$uhrswork, size = n, replace = TRUE)
}
```

```{r summariseall}
# Summarise each sample
meanslength <- dflength %>% 
  select(contains("_")) %>% 
  summarise_all(mean) %>%
  t %>%
  as.data.frame

head(meanslength,10)

# Plot distribution
meanslength %>% ggplot(aes(x=V1)) +
  geom_density(color='black') +
  xlim(20, 70)

bootstraplength <- meanslength %>% summarise_all(funs(mean))
```